---
title: "Exploratory Analysis"
author: "Dima"
date: "`r Sys.Date()`"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(njtPredict)
library(lubridate)
library(ggplot2)
# library(broom)
library(tidyr)
library(knitr)
# devtools::load_all()
 
opts_chunk$set(echo = TRUE,
               message = FALSE,
               prompt = FALSE,
               warning = FALSE,
               cache = TRUE, 
               fig.width = 14,
               fig.height = 8)

```

```{r load_data, cache = TRUE}
data("njt_features")
```

# Overview  
The purpose of this notebook is to characterize & explore the njt data and weather data.  


# Exploring General Attributes  

```{r general_attr}
njt_features %>% 
  count(Line, is_delayed) %>% 
  spread(is_delayed, n) %>%
  mutate(pct_delays = Yes/(No+Yes)*100) %>%
  arrange(desc(pct_delays)) %>%
  kable
```

Not suprisingly, Corridor has the most amount of delays.  

# Exploring Calendar Attributes  

## Delays By Day of Weekday/Line
```{r delay_weekday_line}
njt_features %>% 
  count(Line, dep_wday, is_delayed) %>% 
  spread(is_delayed,n) %>% 
  mutate(pct_delays = Yes/(No+Yes)*100) %>% 
  ggplot(aes(dep_wday, pct_delays)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~Line, scale = "free", ncol = 5) +
  theme_classic()
```

## Delays By Day of Month/Line  
```{r delay_month_line}
njt_features %>% 
  count(Line, dep_mon, is_delayed) %>% 
  spread(is_delayed,n) %>% 
  mutate(pct_delays = Yes/(No+Yes)*100) %>% 
  ggplot(aes(dep_mon, pct_delays)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~Line, scale = "free_y", ncol = 5) +
  theme_classic()
```

## Delays By Day of Hour/Line  
```{r delay_hour_line}
njt_features %>% 
  count(Line, dep_hour, is_delayed) %>% 
  spread(is_delayed,n) %>% 
  mutate(pct_delays = Yes/(No+Yes)*100) %>% 
  ggplot(aes(dep_hour, pct_delays)) + 
  geom_bar(stat="identity") + 
  facet_wrap(~Line, scale = "free_y", ncol = 5) +
  theme_classic()
```

## Time to Last Delay  
### Fraction of delays when previous delay occurred in the last 90 minutes
```{r delay_window}
previous_delay_window = 90
```

#### Using any delay on the line  
```{r time_tolast_line}
njt_features %>% mutate(prdelay = ttl_line <= previous_delay_window) %>% 
  filter(!is.na(prdelay)) %>% 
  count(Line, prdelay, is_delayed) %>%
  spread(is_delayed,n) %>% 
  mutate(pct_delays = Yes/(No+Yes)*100) %>% 
  ggplot(aes(prdelay, pct_delays)) + 
  geom_bar(stat="identity", position = "dodge") + 
  facet_wrap(~Line, scale = "free_y", ncol = 5) +
  theme_classic() +
  xlab(paste("Delay Occurred withing the last",previous_delay_window,"minutes")) +
  ylab("% of all Trains")
```

#### Using delays with fixed departure
```{r time_tolast_linedep}
njt_features %>% mutate(prdelay = ttl_dep_line <= previous_delay_window) %>% 
  filter(!is.na(prdelay)) %>% 
  count(Line, prdelay, is_delayed) %>%
  spread(is_delayed,n) %>% 
  mutate(pct_delays = Yes/(No+Yes)*100) %>% 
  ggplot(aes(prdelay, pct_delays)) + 
  geom_bar(stat="identity", position = "dodge") + 
  facet_wrap(~Line, scale = "free_y", ncol = 5) +
  theme_classic() +
  xlab(paste("Delay Occurred withing the last",previous_delay_window,"minutes")) +
  ylab("% of all Trains")
```

#### Using delays with fixed departure & arrival
```{r time_tolast_linedepari}
njt_features %>% mutate(prdelay = ttl_dep_arv_line <= previous_delay_window) %>% 
  filter(!is.na(prdelay)) %>% 
  count(Line, prdelay, is_delayed) %>%
  spread(is_delayed,n) %>% 
  mutate(pct_delays = Yes/(No+Yes)*100) %>% 
  ggplot(aes(prdelay, pct_delays)) + 
  geom_bar(stat="identity", position = "dodge") + 
  facet_wrap(~Line, scale = "free_y", ncol = 5) +
  theme_classic() +
  xlab(paste("Delay Occurred withing the last",previous_delay_window,"minutes")) +
  ylab("% of all Trains")
```

# Using Weather Data  
```{r}

```


# Comparing Reasons
## NJT vs Amtrak

## Types of Delays
